Portfolios of Subgraph Isomorphism Algorithms

### 2015-12-05 paper/paper.tex
Chrisine's changes

\usepackage{complexity}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{cleveref}                  % no need to type Figure etc

@@

\begin{abstract}
The subgraph isomorphism problem is a computationally challenging problem with
lots of important practical applications, for example in computer vision,
biochemistry, and model checking. There are a number of state-of-the-art
algorithms for solving it, each of which has its own performance
characteristics. As with many other hard problems, the single best choice of
algorithm overall is rarely the best algorithm for each instance. We develop a
portfolio approach that leverages novel features to characterise subgraph
isomorphism problems to dynamically decide which algorithm to use on a
per-instance basis. We demonstrate significant performance improvements on a
large set of hard benchmark problems. In addition, we show how algorithm
selection models can be leveraged to gain new insights into what affects the
performance of an algorithm.
\end{abstract}

@@

The \emph{(non-induced) subgraph isomorphism problem} is to find an injective mapping from a small
\emph{pattern} graph to a large \emph{target} graph which preserves adjacency. This problem is
\NP-complete, but applications in chemistry and computer vision and model checking and some other
things we should cite mean various exact algorithms exist:

\begin{itemize}
    \item VF2 \cite{Cordella:2004} is a backtracking search algorithm which is especially fast on
        trivially satisfiable instances (which are common in certain application areas).
    \item LAD \cite{Solnon:2010} and SND \cite{Audemard:2014} do very strong reasoning, at the
        expense of making few guesses per second when working with larger graphs.
    \item Most recently, the Glasgow algorithm \cite{McCreesh:2015} does an expensive preprocessing
        pass once, followed by weaker reasoning during search.
\end{itemize}

\noindent The experiments by the Glasgow authors indicated that overall, the Glasgow algorithm is
the single best algorithm when evaluated across a wide range of problem instances. However, they
also observed that on an instance by instance basis, it is often not the single best.  Additionally,
the Glasgow algorithm has a parameter, which controls the lengths of paths used when reasoning about
non-adjacent vertices.  The choice of paths of length 3 was used as a reasonable compromise---longer
paths lead to prohibitively expensive preprocessing on larger, denser instances. Again, this is
often not the best choice on an instance by instance basis: sometimes path-based reasoning gives no
benefit at all, sometimes considering only paths of length 2 suffices, occasionally paths of
length 4 are helpful, and even looking at paths of length 3 is relatively expensive on some graphs.

?? Also, we can use supplemental graphs in LAD. Christine: can you explain what you implemented?

We could try to come up with rules to pick an algorithm on an instance by instance basis using our
knowledge of how the algorithms behave: for example, maybe we should use the Glasgow algorithm with
paths of length 2 with large target graphs (but how large?), and only use length 3 on smaller
graphs. Also, we could guess that LAD might do better than Glasgow on smaller, heavily-structured
graphs. However, this is not very systematic.

?? Lars: a paragraph of fluff about instance selection please.

We show that picking algorithms and parameter choices on an instance by instance basis, by
considering features of the pattern and target graphs, gives better overall performance than any
single algorithm. This is the first time this has been done for subgraph isomorphism, and we must
consider an unusual factor: our features must look at pairs of inputs, and the relationships between
them.

?? Lars: do you know of any other problems where you have two inputs rather than one? Or is this
novel?

?? Can we get better results if we explicitly consider, say, how close the target density and
pattern density are?

\section{Experiments}

@@

?? Lars: your stuff here.

\subsection{Problem instances}

All instances are available in a simple text format
\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}.

Real-world and application-derived graphs:

\begin{itemize}
    \item The LV graphs (family 2). Also, the same targets, but with larger pattern graphs (family
        11, Multiple edges have been combined into a single edge.)
    \item CVIU11 images (family 8)
    \item CVIU11 meshes (family 10)
    \item PR15 (family 9)
\end{itemize}

\noindent Randomly generated instances with particular properties:

\begin{itemize}
    \item Bounded-degree random graphs (family 3, 4) \cite{GraphDatabase}
    \item Random 4d mesh graphs (family 5, 6) \cite{GraphDatabase}
    \item Random graphs (family 7) \cite{GraphDatabase}
    \item Randomly generated scale-free graphs. (family 1) \cite{Solnon:2010}
    \item Phase transition (family 12)
\end{itemize}

\noindent Note that this is not like SAT, where we can reasonably split things up based upon which
family they're from.

\subsection{Machine learning voodoo stuff}

\subsection{Results}

?? Can we do better than picking the single best on a family by family basis? Bear in mind that we
don't have the family as a usable feature.

@@

This stuff works respectably well, considering we're dealing with two inputs rather than one.

We have only looked at sequential algorithms: the Glasgow algorithm has a parallel implementation,
and a similar technique could easily be used to parallelise LAD.  We have also not considered the
induced variant of the problem, nor instances involving labels or directed edges.  We could also
look at other encodings. For example, a direct SAT requires $O(v^4)$ clauses, but may be good in
certain situations.

>>>>

\usepackage{complexity}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{cleveref}                  % no need to type Figure etc
\usepackage{xcolor,colortbl}

@@

\begin{abstract}
Subgraph isomorphism is a computationally challenging problem with lots of important practical applications, for example in computer vision, biochemistry, and model checking. There are a number of state-of-the-art algorithms for solving it, each of which has its own performance characteristics. As with many other hard problems, the single best choice of algorithm overall is rarely the best algorithm for each instance. We develop a portfolio approach that leverages novel features to characterise subgraph isomorphism problems to dynamically decide which algorithm to use on a per-instance basis. We demonstrate significant performance improvements on a large set of hard benchmark problems. In addition, we show how algorithm selection models can be leveraged to gain new insights into what affects the performance of an algorithm.\end{abstract}

@@

The subgraph isomorphism problem is to find an injective mapping from vertices of a small \emph{pattern} graph to vertices of a large \emph{target} graph which preserves adjacency. This \NP-complete problem has lots of important practical applications, for example in computer vision, biochemistry, and model checking. There exist various exact algorithms, which have been compared on a large benchmark in  \cite{McCreesh:2015}. These experiments indicated that the single best algorithm depends on the CPU time limit considered: for very small (resp. larger) CPU time limits, VF2 \cite{Cordella:2004} (resp. the Glasgow algorithm \cite{McCreesh:2015}) has better success rates. Furthermore, they also showed us that on an instance by instance basis, other algorithms may become better.

In this paper, we report experimental results on a larger set of 8 algorithms, including new variants, and a larger benchmark. We show that 5 of these algorithms are single best algorithms, depending on the CPU time limit considered, and that combining preprocessings with an algorithm selection on an instance by instance basis allows us to achieve better overall performance than any single algorithm.

\section{Algorithms for subgraph isomorphism}

We consider the following exact algorithms:
 VF2 \cite{Cordella:2004}, which is a simple backtracking search algorithm, especially fast on trivially satisfiable instances;
 LAD \cite{Solnon:2010}, which  does very strong reasoning based on local all different redundant constraints;
 and the Glasgow algorithm \cite{McCreesh:2015}, which does an expensive preprocessing pass once, followed by weaker reasoning during search.
 The Glasgow algorithm has a parameter, which controls the lengths of paths used when reasoning about non-adjacent vertices.  In experiments reported in \cite{McCreesh:2015}, the choice of paths of length 3 was used as a reasonable compromise---longer paths lead to prohibitively expensive preprocessing on larger, denser instances. This is
often not the best choice on an instance by instance basis: sometimes path-based reasoning gives no benefit at all, sometimes considering only paths of length 2 suffices, occasionally paths of length 4 are helpful, and even looking at paths of length 3 is relatively expensive on some graphs. Therefore, in this new study, we consider 4 different instances of the Glasgow algorithm, with path lengths varying from 1 to 4.

We introduce two new variants of LAD. The first variant is a weaker LAD filtering (called LAD preprocessing) which is applied once, without performing a backtracking search, and detects inconsistencies on many instances: For each pattern node $u$, we check that there exists at least one target node $v$ such that for each neighbor $u'$ of $u$ there exists a different neighbor $v'$ of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$. LAD preprocessing is an incomplete algorithm as it can only detect inconsistency of some instances: When it does not detect inconsistency, the instance may either be consistent or not. The time complexity of LAD preprocessing is ${\cal O}(n(n'+p'))$ where $n$ and $n'$ are the number of pattern and target nodes, respectively, and $p$ is the number of target edges.

The second variant of LAD is called new LAD and it is stronger than the default LAD introduced in  \cite{Solnon:2010}. It combines LAD filtering with new invariant properties borrowed from SND \cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of paths of length 2 between $u$ and $v$, and each node $u$ with the number of cycles of length 3 passing through $u$, and to add the constraint that the label of a pattern node (resp. edge) must be smaller than or equal to the label of its associated target node (resp. edge).

\section{Benchmark}

We consider 5725 instances which are available in a simple text format\footnote{\url{http://liris.cnrs.fr/csolnon/SIP.html}}. These instances are grouped into 12 classes.
\begin{itemize}
\item Class 1 contains randomly generated scale-free graphs \cite{}.
\item Classes 2 and 3 contain various kinds of graphs coming from \cite{}: class 2 contains small instances generated from the 50 first graphs of the benchmark, whereas class 3 contains larger instances where pattern graphs are coming from class 2, and target graphs are coming from the next 50 graphs of the benchmark.
\item Classes 4 to 8 contain randomly generated graphs coming from \cite{GraphDatabase}: bounded-degree graphs for classes 3 and 4, regular meshes for classes 5 and 6, and random graphs with uniform edge probabilities for class 7.
\item Classes 9 and 10 contain instances coming from images \cite{}.
\item Class 11 contains instances coming from meshes modeling 3D objects \cite{}.
\item Class 12 contains random graph instances belonging to the phase transition.
\end{itemize}
Note that Classes 3 and 12 were not considered in the previous experimental study reported in \cite{McCreesh:2015}.

\section{Experimental comparison}

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{|c||r||r|r|r||r|r|r|r||r|r|}
\hline
$t$ & VF2 & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c||}{Glasgow}& VBS & $\delta$\\
&&pre&def&new&1&2&3&4&&\\\hline
$10^0$ & 300 &  \cellcolor{blue!25}1126 & 661 & 599 & 386 & 95 & 39 & 12 & 1132 & 6\\\hline
$10^1$ & 1699 & 1858 & 2101 &  \cellcolor{blue!25}2287 & 2061 & 1215 & 455 & 208 & 3388 & 101\\\hline
$10^2$ & 2834 & 1912 & 3376 & 3708 &  \cellcolor{blue!25}3943 & 3344 & 2295 & 1329 & 4636 & 693\\\hline
$10^3$ & 3458 & 1918 & 4233 & 4531 &  \cellcolor{blue!25}4908 & 4763 & 4262 & 3426 & 5170 & 262\\\hline
$10^4$ & 3701 & 1919 & 4888 & 5025 & 5156 &  \cellcolor{blue!25}5253 & 5022 & 4408 & 5332 & 79\\\hline
$10^5$ & 3850 & 1919 & 5108 & 5193 & 5301 &  \cellcolor{blue!25}5377 & 5292 & 5082 & 5434 & 57\\\hline
$10^6$ & 3977 & 1919 & 5244 & 5307 & 5386 &  \cellcolor{blue!25}5452 & 5451 & 5235 & 5500 & 48\\\hline
$10^7$ & 4082 & 1919 & 5336 & 5414 & 5458 & 5517 &  \cellcolor{blue!25}5518 & 5425 & 5567 & 49\\\hline
$10^8$ & 4191 & 1919 & 5423 & 5479 & 5508 &  \cellcolor{blue!25}5561 & 5560 & 5554 & 5608 & 47\\\hline
\end{tabular}
\end{center}
\caption{Number of solved instances at different CPU time limits: Each line displays a time limit $t$ (in milliseconds) followed by the number of instances solved within this time limit by VF2, LAD preprocessing, default LAD, and new LAD, Glasgow with  the lengths of paths limited to 1, 2, 3 and 4, and the VBS. The best single algorithm is highlighted in blue, and the difference between VBS and single best if displayed in the last column ($\delta$).\label{expTime}}
\end{table}

Table \ref{expTime} displays the number of instances solved at different CPU time limits, ranging from 1 to $10^8$ milliseconds. It shows us that the best single solver depends on the time limit considered. Simple approaches like LAD preprocessing and VF2 are able to solve easy instances very quickly, in a few milliseconds. However, they are not able to solve harder instances: LAD preprocessing is an incomplete approach which can only detect rather trivial unconsistencies; VF2 performs a basic backtracking search and is very fast on easy instances, because it does not compute expensive invariants. Default LAD, new LAD, and the Glasgow algorithms solve more instances than VF2 and LAD preprocessing when considering longer time limits: 10 ms for default and new LAD and Glasgow 1, and 100 (resp. 1000 and 10000) ms for Glasgow 2 (resp. 3 and 4). The Virtual Best Solver (VBS), which considers the best algorithm for each instance separately, obtains much better results, showing us that the algorithms have complementary performance. The difference between VBS and single best (column $\delta$ of Table \ref{expTime}) is more particularly important for rather small CPU time limits. In particular, VBS solves 693 more instances than the single best (Glasgow 1) when the limit is 100 ms. In many applications, it is important to have the fastest possible algorithm. For example, in pattern recognition applications, we often have to solve subgraph isomorphism problems for a very large number of graphs (in order to find a pattern image in a large database of target images, for example), so that having an algorithm that is able to solve an instance in 100 ms instead of 1000 ms makes a big difference. Therefore, it is important to select the best algorithm for each instance.

Table \ref{expClass} shows us that we cannot reasonably split things up based upon which class instances are comin from, unlike SAT. For all classes, there are always at least 2 algorithms which are the best for at least one instance of the class. In particular, for classes 2 and 3, each algorithm is the best for at least one instance (except Glasgow 4 for Class 2).

\begin{table}
\begin{center}
\begin{tabular}{|c||r||r|r|r||r|r|r|r|}
\hline
Class & VF2 & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c|}{Glasgow}\\
&&pre&def&new&1&2&3&4\\\hline
1\\\hline
2\\\hline
3\\\hline
4\\\hline
5\\\hline
6\\\hline
7\\\hline
8\\\hline
9\\\hline
10\\\hline
11\\\hline
12\\\hline

\end{tabular}
\end{center}
\caption{Number of times each algorithm is best, for each class.\label{expClass}}
\end{table}

\section{Algorithm selection approach}

Describe features + selection approach.

Table 3 shows the performance of our algorithm selection approach, compared to the VBS (which is the upper bound of what an algorithm selection approach can achieve), and the single best solver, at different CPU time limits (the single best corresponds to results highlighted in blue in Table 1). Comment Table 3: Mean runtime but also number of solved instances at different CPU time limits. Logically, LLAMA may solve less instances than single best for small time limits because it spends time to compute features and choose a solver, but it should be better for larger time limits...

\begin{table}
\begin{tabular}{|l|r|rrrrrrrrr|}
&mean & \multicolumn{9}{c|}{Number of solved instances}\\
&runtime & $10^0$ &  $10^1$ &  $10^2$ &  $10^3$ &  $10^4$ &  $10^5$ &  $10^6$ &  $10^7$ &  $10^8$\\\hline
VBS & 2375.9 & 1132 & 3388 & 4636 & 5170 & 5332 & 5434 & 5500 & 5567 & 5608\\\hline
Single best & 3177.6 & 1126 & 2287 & 3943 & 4908 & 5253 & 5377 & 5452 & 5518 & 5561\\\hline
LLAMA & 2706.3\\\hline
\end{tabular}
\caption{}
\end{table}

Add Figure 1 which compares single best at $10^8$ ms (Glasgow 2) with LLAMA ?

@@
